A Mechanised Library for Generic (Co)Programming

### 2024-01-21 paper/paper.tex

As mentioned, some functions, though well-defined, cannot be accepted by the
proof assistant. One example of this fact is the quicksort algorithm given by
the \haskell{partition} and \haskell{quicksort} function below (written in
Haskell code):
\begin{minted}{haskell}
partition :: [a] -> ([a], a, [a])
partition (x:xs) = (filter (x <=) xs, x, filter (x >) xs)

@@

Since then this result has had huge impact in the functional programming
community for structuring functional programs in a way that it guarantees
termination (or productivity) in programming languages with general
recursion~\cite{DanielssonHJG06}.

@@

The first option, helps the programmer but does not guarantee termination, since
the programmer can still feed to the recursion scheme a functor that does not
have a fixed-point or that has an algebra that does not terminate.

The second option forbids non-termination, but does not allow all terminating
programs.

With this work we intend to begin a study of the applications of recursion
schemes that goes beyond functional programming, that is to dependent type
theories such as a the ones employed in proof assistants like Agda or Coq where
non-termination is outright rejected.  This has been a long-standing challenge
to come up with a sound algorithm which rejects all programs that do not
terminate (soundness) and accepts as many well-defined programs as possible
(completeness). Our aim is to provide an additional tool to the every-day
programmer to be able to extend the range of recursive programs we can write.

>>>>

As mentioned, some functions, though well-defined, cannot be accepted by the
pThis specific module poses a considerable challenge for the current student cohort. However, upon evaluating their performance in assignments, I can affirm that the majority are managing well. Recognizing the need for change in the teaching methodology to foster increased engagement, I am considering shifting from traditional slides to utilising my iPad/whiteboard during lectures. Additionally, incorporating in-class group exercises aims to slow down the lecture pace and encourage active participation

The classic whiteboard teaching approach is known to enhance student involvement by compelling the lecturer to moderate their speed, allowing students to absorb the material more gradually. Supplementary materials will be provided to support their learning outside the classroom.
I intend to meticulously gauge the impact of these modifications through assignment evaluations and ongoing observation of student participation. Any deviations from the expected outcomes, indicated by a lack of in-class responses, will be promptly addressed. Encouragingly, I have already shared these proposed changes with my peers in the PGCHE, receiving positive feedback and support for the initiative.

I've already shared some of my plans with the PGCHE instructors, and the response has been quite positive. I will continue to stay in touch with them for further feedback on future changes I intend to make.
roof assistant. One example is the quicksort algorithm. This is given by the \haskell{partition} and \haskell{quicksort} function below (written in
Haskell code):
\begin{minted}{haskell}
partition :: [a] -> ([a], a, [a])
partition (x:xs) = (filter (x <=) xs, x, filter (x >) xs)

@@

This result has given a much needed unifying view of recursion scheme which
helped understanding their nature and also in providing a unified way of
reasoning about recursive programs by means of fusion laws.

@@

The type theoretical approach instead is to reject all definitions which cannot
be proven well-defined. It is safer, but incomplete.

With this work we intend to begin a study of the applications of recursion
schemes that goes beyond functional programming, that is to dependent type
theories such as a the ones employed in proof assistants like Agda or Coq where
non-termination is outright rejected.

This has been a long-standing challenge to come up with a sound algorithm which
rejects all programs that do not terminate (soundness) and accepts as many
well-defined programs as possible (completeness). Our aim is to provide an
additional tool to the every-day programmer to be able to extend the range of
recursive programs we can write.
